Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-11-05 Cost_manuscript.Rmd
moving oc around

  This experiment probed whether the value-modulatory effects of each type of demand depended on the ability to obtain rewards through other behavioral means. If the cost or value of a demand are intrinsic, the relative preferences displayed above should be robust to alternatives. The experimental setup was mostly preserved, with 3 key differences. First, all participants experienced cognitive effort, physical effort, and delay. Second, each block contained offers with one type of effort and passive waiting (allowing for comparisons in acceptances through the common wait condition). We refer to observations from each block's wait trials by their effort pair (i.e. "Wait-C" for cognitive, and "Wait-P" for physical). Finally, in this experiment the handling and travel times were fixed at 10 s and 6 s, respectively. We therefore compared choices from the within-subjects experiment to the matching time condition from the between-subjects experiment. We hypothesized that reward decision patterns would match those of the previous experiment, and that interleaving effort and delay would produce persistent preferences for passive delay trials.

@@

  Analyses from the previous experiment were repurposed to examine cost differences within individuals. A mixed-effects logistic regression once again showed that larger reward amounts significantly increased acceptance probabilities ($\beta$ = `r historyCoeffs_wth["Offer", "Estimate"]`, SE = `r historyCoeffs_wth["Offer", "Std. Error"]`, *p* `r report_p(historyCoeffs_wth["Offer", "Pr(>|z|)"])`), and that participants became more selective if they had recently observed large offers ($\beta$ = `r historyCoeffs_wth["recentOffers", "Estimate"]`, SE = `r historyCoeffs_wth["recentOffers", "Std. Error"]`, *p* `r report_p(historyCoeffs_wth["recentOffers", "Pr(>|z|)"])`). Model comparisons suggested that participants were influenced by the `r historyIndex_wth + 1` most recent rewards (AIC = `r AIC(historySize_wth[[historyIndex_wth]])`. However, this time participants were unaffected by recent reward misses (*p* > 0.05).
  As before, a linear model predicting earnings showed that participants who completed too few or too many trials earned the least, in line with the experimental design (Figure 8A; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, Beta = `r round(summary(lmEarn)$coefficients[3,1], digits= 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, R-squared = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). When assessing the optimality of acceptances for 4 and 8 cents (99% of 20 cent offers were accepted), chi-squared tests against the optimal rate of 50% showed that participants under-accepted physical effort trials (proportion accepted = `r round(summary_costvsreward["Physical", "probability"], digits = 2)`, $\chi^2$ = `r round(summary_costvsreward["Physical", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Physical", "pval"])`; all other *p* > 0.05). Despite this deviation, participants earned a similar monetary amount per cost on average (mean Cognitive = `r filter(earnings_within, Cost == "Cognitive")$mEarn`, SD = `r filter(earnings_within, Cost == "Cognitive")$sdEarn`; mean Wait-C = `r filter(earnings_within, Cost == "Wait-C")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait-C")$sdEarn`; mean Physical = `r filter(earnings_within, Cost == "Physical")$mEarn`, SD = `r filter(earnings_within, Cost == "Physical")$sdEarn`; mean Wait-P = `r filter(earnings_within, Cost == "Wait-P")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait-P")$sdEarn`). These results convey that participants in both experiments were similarly influenced by the characteristics of the environment, although participants in the present experiment gravitated more towards optimality (**caveat: the same percentage can be produced by many non-optimal combinations, including accepting half of each reward. So neither approach really gets at this too well.**).

@@

### **Within-subjects: Modeling the subjective opportunity cost in each condition**

@@

@@

**Possibilities**

@@

  Together, these findings suggest that 1) foraging paradigms can provide a reference point to evaluate how different demands modulate value; 2) a simple OC model can approximate people's behavior in such single-option, take it or leave it choices; and 3) when foraging in environments of equivalent richness, cognifitvely effortful demands can increase the attractiveness of offers rather than discount them (unlike other forms of cost). In the next experiment we explored whether this apparent value was inherent or malleable.
## Discussion

>>>>

  If the cost or value of a demand were intrinsic, the relative preferences displayed so far should be robust to the ability to obtain the same rewards through different demands. We investigated this possibility by modifying the foraging task, such that every participant experienced every type of demand in blocks that interleaved wait trials with either physical or cognitive effort (fixed at 10 s handling and 6 s travel times, to allow for comparisons between experiments). We refer to observations from each block's wait trials by their effort pair (i.e. "Wait-C" for cognitive, and "Wait-P" for physical). We hypothesized that reward decision patterns would match those of the previous experiment, and that interleaving effort and delay would produce persistent preferences for wait trials.

@@

  Analyses from the previous experiment were repurposed to examine cost differences within individuals. A mixed-effects logistic regression once again showed that larger reward amounts significantly increased acceptance probabilities ($\beta$ = `r historyCoeffs_wth["Offer", "Estimate"]`, SE = `r historyCoeffs_wth["Offer", "Std. Error"]`, *p* `r report_p(historyCoeffs_wth["Offer", "Pr(>|z|)"])`), and that participants became more selective if they had recently observed large offers ($\beta$ = `r historyCoeffs_wth["recentOffers", "Estimate"]`, SE = `r historyCoeffs_wth["recentOffers", "Std. Error"]`, *p* `r report_p(historyCoeffs_wth["recentOffers", "Pr(>|z|)"])`). Model comparisons suggested that participants were influenced by the `r historyIndex_wth + 1` most recent rewards (AIC = `r AIC(historySize_wth[[historyIndex_wth]])`. However, this time participants were unaffected by recent reward misses (*p* > 0.05).
  As before, a linear model predicting earnings showed that participants who completed too few or too many trials earned the least, in line with the experimental design (Figure 8A; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, Beta = `r round(summary(lmEarn)$coefficients[3,1], digits= 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, R-squared = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). When assessing the optimality of acceptances for 4 and 8 cents (99% of 20 cent offers were accepted), chi-squared tests against the optimal rate of 50% showed that participants under-accepted physical effort trials (proportion accepted = `r round(summary_costvsreward["Physical", "probability"], digits = 2)`, $\chi^2$ = `r round(summary_costvsreward["Physical", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Physical", "pval"])`; all other *p* > 0.05). Despite this deviation, participants earned a similar monetary amount per cost on average (mean Cognitive = `r filter(earnings_within, Cost == "Cognitive")$mEarn`, SD = `r filter(earnings_within, Cost == "Cognitive")$sdEarn`; mean Wait-C = `r filter(earnings_within, Cost == "Wait-C")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait-C")$sdEarn`; mean Physical = `r filter(earnings_within, Cost == "Physical")$mEarn`, SD = `r filter(earnings_within, Cost == "Physical")$sdEarn`; mean Wait-P = `r filter(earnings_within, Cost == "Wait-P")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait-P")$sdEarn`). These results convey that participants in both experiments were similarly influenced by the characteristics of the environment, although participants in the present experiment gravitated more towards optimality.

@@

### **Consistency and stability**

@@

  Together, these findings suggest that 1) foraging paradigms can provide a reference point to evaluate how different demands modulate value; 2) a simple OC model can approximate people's behavior in such single-option, take it or leave it choices; and 3) when foraging in environments of equivalent richness, cognifitvely effortful demands can increase the attractiveness of offers rather than discount them (unlike other forms of cost). In the next experiment we explored whether this apparent value was inherent or malleable.

@@

@@

## Discussion
